---
title: "R tutorial and notes"
author: "GD"
date: "7/20/2023"
output: html_document
---
# Helpful packages
```{r}
# General ones to always have
library(dplyr)
library(tidyverse)

# Reading in and exporting files
library(readxl)
library(openxlsx)

# Data visualization 
library(ggplot2)
library(ggpubr)

# Making tables
library(finalfit)
library(table1)
library(flextable)
library(gridExtra)

# Other helpful packages
library(psych)
library(Hmisc)
library(jtools)
```
# Merging datafames
## Only add in specific columns
```{r}
# Merge specific columns of interest based on a column shared by the two DFs
df_to_update <- df_to_update %>% 
  left_join(df_with_new_cols %>%
              select(shared_column, new_col_1, new_col_2),
            by="shared_column")
```
## Cleanly merge many dataframes
```{r}
# Create a list variable of the dataframes of interest
list_df = list(df1,df2,df3,df4)

# Merge DFs
merged_df <- bind_rows(list_df) %>%
  # If you want to arrange by two specific columns together, add this line 
  # (in this example case, we are getting IDs in number, timepoint, and rescan order) (122-T3, 122x-T3, 122-T4, 123-T1, etc.)
  arrange(Subject_ID_BIDS, Timepoint_BIDS)
```
# DF manipulation
## Subsetting
```{r}
# Create subsets of males at T2
subset_df <- original_df %>% 
  filter(Sex == 1,
         Timepoint == "T2")
```

# Make visual changes to dataframes
## Relocate a column
```{r}
df <- df %>% 
  relocate(column_youre_moving, .after = column_you_want_on_the_left)
```
## Remove columns you don't want
```{r}
df2 <- df1 %>%
  select(-(c('column_1','column_2')))

# If you want to check the columns this removed
## To get the columns in 'df1' that are not present in 'df2':
columns_only_in_df1 <- setdiff(names(df1), names(df2))
## To get the columns in 'df2' that are not present in 'df1':
columns_only_in_df2 <- setdiff(names(df2), names(df1))
## Print the results
print(columns_only_in_df1)
print(columns_only_in_df2)

```
## Rename columns
```{r}
df <- df %>% 
  rename(new = old)
```

# Sanity checks
## Check what rows are in one df but not the other
```{r}
rows_not_in_cleaned <- anti_join(df_original, df_new)
```
## Check what columns are in one df but not the other
```{r}
columns_only_in_df1 <- setdiff(names(df1), names(df2))
# Or, see what columns are present in both df1 and df2 
common_columns <- intersect(colnames(df1), colnames(df2))
```
# Console tips
- rm(df): removes only specific elements from your environment
- describe(df): gives statistical info about df
- summary(df): gives statistical info about each column in df
- glimpse(df): gives column names and cell types, lets you preview the first few rows
- colnames(df): lists the column names in df

# Statistics
## Correlation coefficients
```{r}
cor.test()
# --> base R test to calculate the Pearson correlation coefficient between two numeric vectors
# --> returns correlation coefficient, p-value, and (optionally) confidence intervals 
corr.test()
# --> psych package test to calculate correlation coefficients for multiple variables 
# --> can take in data frame or matrix, etc.
```
