---
title: "R tutorial and notes"
author: "GD"
date: "7/20/2023"
output: html_document
---
# Helpful packages
```{r}
# General ones to always have
library(dplyr)
library(tidyverse)

# Reading in and exporting files
library(readxl)
library(openxlsx)

# Data visualization 
library(ggplot2)
library(ggpubr)

# Making tables
library(finalfit)
library(table1)
library(flextable)
library(gridExtra)

# Other helpful packages
library(psych)
library(Hmisc)
library(jtools)
```
# Mac shortcuts
    - commmand option i: inserts a chunk
    - command shift m: inserts a pipe
# Console tips
** visual alterations **
    - rm(df): removes only specific elements from your environment
** understanding your df as a whole **
    - describe(df): gives statistical info about df
    - summary(df): gives statistical info about each column in df
    - glimpse(df): gives col names, data class types, previewsfirst few rows
    - colnames(df): lists the column names in df
** understanding specifics **
    - typeof(): determines the type or storage mode of any object 
    - class(): determines what class attribute the object inherited
    - length(): determines length of an R object

# Random R tips
## Clean the environment
```{r clean}
rm(list=ls(all=TRUE)) 
```
## Console tips
    ?function() gives you information on how to use that function

# Reading in data
```{r}
read_excel()
read.csv()

# can use pattern to read in files ending in same pattern in same folder
data_pathnames <- list.files(
  path = "dir path",
  pattern = "*.csv", full.names = TRUE)

```

# Make visual changes to dataframes
## Relocate a column
```{r}
df <- df %>% 
  relocate(column_youre_moving, .after = column_you_want_on_the_left)
```
## Remove columns you don't want
```{r}
df2 <- df1 %>%
  select(-(c('column_1','column_2')))

# If you want to check the columns this removed
## To get the columns in 'df1' that are not present in 'df2':
columns_only_in_df1 <- setdiff(names(df1), names(df2))
## To get the columns in 'df2' that are not present in 'df1':
columns_only_in_df2 <- setdiff(names(df2), names(df1))
## Print the results
print(columns_only_in_df1)
print(columns_only_in_df2)

```
## Rename columns
```{r}
df <- df %>% 
  rename(new = old)
```

# Data wrangling
```{r}
count()
# Shows the counts of all the distinct values in a column
# Can do grouped counts by putting two variables in the columns 

select() 
# Allows you to select a subset of the columns in the data frame 
# Put a ‘-’ before the variable to remove that variable from the df
# Example: select(column_name_1, column_name_2)
# If you have a lot of variables you are selecting for, use c() to ‘count’ -- this is a cleaner way to list a lot of different variables (*note: the order you put them in is the order these variables will appear in the new df*)

Indexing
# Instead of using select(), can use hard brackets directly to ‘index’ through dataframe: df[#,#]
# The first index refers to the row, the second to the column

slice()
# Selects out rows based on their row number

filter()
# Lets you apply logical (and other) operators to subset the data and select rows based off specific values
# Ex: filter(Gender == "Female")

```
## Altering variables
```{r}
# Add new variables but preserve existing ones
mutate()

# Add new variables but drop existing ones
transmute()

# Replace certain text
str_replace(variable)
# to replace everything coming before a pattern you want to keep the same
# //(x-T4)”

```

## Subsetting dfs
```{r}
# Create subsets of males at T2
subset_df <- original_df %>%
  filter(Sex == 1,
         Timepoint == "T2")
```
## Easily viewing DFs
```{r}
sample_n(df, 3) #insert number of rows you want to preview 
```

# Merging datafames
## Merge full dfs
```{r}
left_join()
full_join()
r_bind()
c_bind()
```

## Add in specific columns
```{r}
# Merge specific columns of interest based on a column shared by the two DFs
df_to_update <- df_to_update %>% 
  left_join(df_with_new_cols %>%
              select(shared_column, new_col_1, new_col_2),
            by="shared_column")
```
## Cleanly merge many dataframes
```{r}
# Create a list variable of the dataframes of interest
list_df = list(df1,df2,df3,df4)

# Merge DFs
merged_df <- bind_rows(list_df) %>%
  # If you want to arrange by two specific columns together, add this line 
  # (in this example case, we are getting IDs in number, timepoint, and rescan order) (122-T3, 122x-T3, 122-T4, 123-T1, etc.)
  arrange(Subject_ID_BIDS, Timepoint_BIDS)
```
# DF sanity checks
## Check what rows are in one df but not the other
```{r}
rows_not_in_cleaned <- anti_join(df_original, df_new)
```
## Check what columns are in one df but not the other
```{r}
columns_only_in_df1 <- setdiff(names(df1), names(df2))
# Or, see what columns are present in both df1 and df2 
common_columns <- intersect(colnames(df1), colnames(df2))
```
# Statistics
## Correlation coefficients
```{r}
cor.test()
# --> base R test to calculate the Pearson correlation coefficient between two numeric vectors
# --> returns correlation coefficient, p-value, and (optionally) confidence intervals 
corr.test()
# --> psych package test to calculate correlation coefficients for multiple variables 
# --> can take in data frame or matrix, etc.
```
# Write out DF
```{r}
write.csv(df_variable,
          "/path.csv",
          row.names = FALSE)
```

# Formatting a PDF using Rmd
## Equations
```{r}
# Use $ for latex equations -- examples below:
```
$$
2 = {\lambda}^x
$$
$$
\log 2 = \log {\lambda}^x = x \cdot \log{\lambda}
$$

